---
title: "Capstone Project"
author: "John Plaxton"
date: "November 6, 2018"
bibliography: ['citations.bib']
biblio-style: "apalike"
link-citations: true
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quanteda)
library(readtext)
library(spacyr)
library(tidyverse)
quanteda_options(threads=parallel::detectCores() - 1)
```

## Create a Corpus

Three copora files are provided for the capstone project.  How do I read into R?  The following code answers the questions for the week 1 quiz.

```{r quiz 1 output}
con <- file("./final/en_US/en_US.twitter.txt", "r") 
size<-file.info("./final/en_US/en_US.blogs.txt")
kb<-size$size/1024
mb<-kb/1024
mb
close(con)

twitter <- readLines(con <- file("./final/en_US/en_US.twitter.txt"), encoding = "UTF-8", skipNul = TRUE)
length(twitter)


#Blogs file
blogs<-file("./final/en_US/en_US.blogs.txt","r")
blogs_lines<-readLines(blogs)
close(blogs)
summary(nchar(blogs_lines))

# News file
news<-file("./final/en_US/en_US.news.txt","r")
news_lines<-readLines(news)
close(news)
summary(nchar(news_lines))

# Twitter file
twitter<-file("./final/en_US/en_US.twitter.txt","r")
twitter_lines<-readLines(twitter)
close(twitter)
summary(nchar(twitter_lines))

love<-length(grep("love", twitter_lines))
hate<-length(grep("hate", twitter_lines))
love/hate

grep("biostats", twitter_lines, value = T)

grep("A computer once beat me at chess, but it was no match for me at kickboxing", twitter_lines)


con <- file("./final/en_US/en_US.twitter.txt", "r") 
readLines(con, 1) ## Read the first line of text 
readLines(con, 1) ## Read the next line of text 
readLines(con, 5) ## Read in the next 5 lines of text 
close(con) ## It's important to close the connection when you are done

```

## Create a Corpus

Create a Corpus from the provided .txt files...

Translating unstructured content from a corpus of information into a meaningful knowledge base is the task of NLP.  It is important to determine what information to keep and how to look for patterns in the structure of that information to distill meaning and context.


```{r create_corpus, cache=TRUE}
textin <- readtext("./final/en_US/*.txt", cache = FALSE)
my_corpus <- corpus(textin)
metadoc(my_corpus, "language") <- "english"
summary(my_corpus, showmeta = TRUE)

```
## Exploratory Analysis

The first step in building a predictive model for text is understanding the distribution and relationship between the words, tokens, and phrases in the text. The goal of this task is to understand the basic relationships you observe in the data and prepare to build your first linguistic models.

> We thus define the tidy text format as being a table with one-token-per-row. A token is a meaningful unit of text, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens. This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix. For tidy text mining, the token that is stored in each row is most often a single word, but can also be an n-gram, sentence, or paragraph. In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format. [@TextMining_R]

A token is a meaningful unit of text, most often a word, that we are interested in using for further analysis, and tokenization is the process of splitting text into tokens.

###Tasks to accomplish

1. **Exploratory analysis** - perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora.
2. **Understand frequencies of words and word pairs** - build figures and tables to understand variation in the frequencies of words and word pairs in the data.

###Questions to consider

1. Some words are more frequent than others - what are the distributions of word frequencies?
```{r tokenize}
#tokens(my_corpus, remove_numbers = TRUE, remove_punct = TRUE, remove_separators = TRUE)

# make a dfm, removing stopwords and applying stemming
myStemMat <- dfm(my_corpus, remove = stopwords("english"), stem = TRUE, remove_punct = TRUE)
topfeatures(myStemMat, 20)  # 20 top words
set.seed(100)
textplot_wordcloud(myStemMat, min_count = 6, random_order = FALSE,
                   rotation = .25, 
                   color = RColorBrewer::brewer.pal(8,"Dark2"))
```

```{r tidytext_method}
library(janeaustenr)
library(dplyr)
library(stringr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()

original_books
library(tidytext)
tidy_books <- original_books %>%
  unnest_tokens(word, text)

tidy_booksdata(stop_words)

tidy_books <- tidy_books %>%
  anti_join(stop_words)
tidy_books %>%
  count(word, sort = TRUE)
library(ggplot2)

tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 600) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

2. What are the frequencies of 2-grams and 3-grams in the dataset?
3. How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%?
4. How do you evaluate how many of the words come from foreign languages?
5. Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words in the dictionary to cover the same number of phrases?

```{r cleanup, include=FALSE}
quanteda_options(threads=2)
```